---
title: "TidyFriday"
output: html_notebook
---

# Popecol data viz using TidyTuesday datasets

[Google doc link](https://docs.google.com/document/d/1XWlxHSOURZ4dniOsA1lyoHDCHYhZrx5nvCuxFwkkcTA/edit#heading=h.nj23sjpj5u97)

## Loading packages

```{r}
library(tidyverse)
```

# Dog breeds dataset

Download data to RStudio environment:

```{r}
breed_traits = readr::read_csv ('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_traits.csv')
trait_description = readr::read_csv ('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/trait_description.csv')
breed_rank_all = readr::read_csv ('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_rank.csv')
```

## Spider/Radar plots

Plotting favorite dogs by country (Germany and Brazil) and their personality traits.

Selecting and renaming traits, and removing spaces from breed names.

```{r}
breed_traits <- breed_traits %>% 
  select(Breed, 
         `Affectionate With Family`, 
         `Openness To Strangers`, 
         `Good With Other Dogs`, 
         `Playfulness Level`, 
         `Trainability Level`,
         `Energy Level`, 
         `Watchdog/Protective Nature`, 
         `Barking Level`,
         `Mental Stimulation Needs`,
         `Shedding Level`,
         `Coat Grooming Frequency`) %>% 
  rename("Family affection" = "Affectionate With Family",
         "Stranger friendliness" = "Openness To Strangers",
         "Dog friendliness" = "Good With Other Dogs",
         "Playfulness" = "Playfulness Level",
         "Trainability" = "Trainability Level",
         "Energy" = "Energy Level",
         "Protectivity" = "Watchdog/Protective Nature",
         "Barking" = "Barking Level",
         "Play demand" = "Mental Stimulation Needs",
         "Shedding" = "Shedding Level",
         "Grooming demand" = "Coat Grooming Frequency") %>%
  mutate(Breed=str_squish(Breed))
```

Select top 10 favorite breeds in and turn breed into factor.

```{r}
german_favs <- breed_traits %>% 
  filter(Breed %in% c("Pomeranians", "Boxers", "German Shepherd Dogs", "Rottweilers",
                      "Standard Schnauzers", "Miniature Pinschers", "Dachshunds", 
                      "Great Danes", "Doberman Pinschers", "Weimaraners")) %>% 
  mutate(Breed = factor(Breed))


brazilian_favs <- breed_traits %>% 
  filter(Breed %in% c("Shih Tzu", "Yorkshire Terriers", "Poodles", "Lhasa Apsos",
                      "Bulldogs", "Dachshunds", "Maltese", 
                      "Finnish Spitz", "Miniature Pinschers", "Standard Schnauzers"))%>% 
  mutate(Breed = factor(Breed))  
```

## Radar plot using ggradar

Install/load ggradar package.

```{r}
devtools::install_github("ricardo-bion/ggradar", dependencies = TRUE)
library(ggradar)
```

Plot of favorite dogs in Germany

```{r}
german_favs %>% 
ggradar(grid.min = 1,                    #minumum trait value
          grid.max = 5,                  #maximum trait value
          values.radar = c(1, 5),        
          label.gridline.mid = FALSE,
          axis.label.size = 2,           # trait font size
          axis.label.offset = 1.1,       # trait name distance from the circle
          group.point.size = 2,          # trait value point size
          group.colours = "black",       # line color
          gridline.mid.colour = "gray",  # color of the central circle 
          group.line.width = 0.5,        # line thickness
          grid.label.size = 0) +         # remove label
  facet_wrap(vars(group)) +              # separate plots by breed
  theme(legend.position = "none",        # remove legend
        strip.text = element_text(size = 10)) # breed name font size
```

Plot of favorite dogs in Brazil

```{r}
brazilian_favs %>%  
ggradar(grid.min = 1,                    #minumum trait value
          grid.max = 5,                  #maximum trait value
          values.radar = c(1, 5),        
          label.gridline.mid = FALSE,
          axis.label.size = 2,           # trait font size
          axis.label.offset = 1.1,       # trait name distance from the circle
          group.point.size = 2,          # trait value point size
          group.colours = "black",       # line color
          gridline.mid.colour = "gray",  # color of the central circle 
          group.line.width = 0.5,        # line thickness
          grid.label.size = 0) +         # remove label
  facet_wrap(vars(group)) +              # separate plots by breed
  theme(legend.position = "none",        # remove legend
        strip.text = element_text(size = 10)) # breed name font size
```

#### All options in ggradar

-   ggradar(plot.data, base.size = 15,

-   font.radar = "sans",

-   values.radar = c("0%", "50%", "100%"),

-   axis.labels = colnames(plot.data)[-1],

-   grid.min = 0,

-   grid.mid = 0.5,

-   grid.max = 1,

-   centre.y = grid.min - ((1/9) *(grid.max - grid.min)),*

-   *plot.extent.x.sf = 1,*

-   *plot.extent.y.sf = 1.2,*

-   *x.centre.range = 0.02* (grid.max - centre.y),

-   label.centre.y = FALSE,

-   grid.line.width = 0.5,

-   gridline.min.linetype = "longdash",

-   gridline.mid.linetype = "longdash",

-   gridline.max.linetype = "longdash",

-   gridline.min.colour = "grey",

-   gridline.mid.colour = "#007A87",

-   gridline.max.colour = "grey",

-   grid.label.size = 6,

-   gridline.label.offset = -0.1 \* (grid.max - centre.y),

-   label.gridline.min = TRUE,

-   label.gridline.mid = TRUE,

-   label.gridline.max = TRUE,

-   axis.label.offset = 1.15,

-   axis.label.size = 5,

-   axis.line.colour = "grey",

-   group.line.width = 1.5,

-   group.point.size = 6,

-   group.colours = NULL,

-   background.circle.colour = "#D7D6D1",

-   background.circle.transparency = 0.2,

-   plot.legend = if (nrow(plot.data) > 1) TRUE else FALSE,

-   legend.title = "",

-   plot.title = "",

-   legend.text.size = 14,

-   legend.position = "left")

## Circular barplot with ggplot2 

Rotate table.

```{r}
german_favs_long <- german_favs %>%
  pivot_longer(cols=2:12, names_to ="Trait", values_to="Value")
```

Plot

```{r}
ggplot(german_favs_long) +
  geom_col(aes(x = Trait, y = Value, fill = Trait),
    position = "dodge2",
    show.legend = TRUE,
    alpha = .9) +
  
  # Scale y axis so bars don't start in the center
  scale_y_continuous(
    limits = c(-2, 5),
    expand = c(0, 0),
    breaks = c(1, 3, 5) ) + 
    
  
#Separate by breed
    facet_wrap(~Breed)+
  theme_bw()+


# Add y value texts
annotate(
    x = 0.5, 
    y = 1.2, 
    label = "1", 
    geom = "text", 
    size = 3) +
   
  annotate(
    x = 0.5, 
    y = 3.2, 
    label = "3", 
    geom = "text", 
    size = 3) +
  annotate(
    x = 0.5, 
    y =5.2, 
    label = "5", 
    geom = "text", 
    size = 3) +

# Scale y axis so bars don't start in the center
  scale_y_continuous(
    limits = c(-2, 6),
    expand = c(0, 0),
    breaks = c(1, 3, 5)  ) + 
  
# Make it circular
  coord_polar()+
  
  #Remove axis titles
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.y = element_blank())

```

# Board games dataset

Download data from Github.
```{r}
game_ratings = readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-25/ratings.csv')
game_details = readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-25/details.csv')
```

Create a dataset joining some columns from ratings and details.
```{r}

games = left_join ( 
# select columns (id, name, average) from dataframe “game_ratings”  
  select(game_ratings, id, name, average),  

# select columns (id, boardgamecategory) from dataframe “details”
  select(game_details,id,boardgamecategory),

# variable that is the same in both dataframes
  by = "id") 	%>%				

# Edit categories names
  mutate(boardgamecategory = str_match_all(boardgamecategory,"\'([a-zA-Z\\s]+)\'"))%>%

# Unnest games with more than one category
  unnest(boardgamecategory) %>%

# Remove the description column of boardgamecategory
  mutate(boardgamecategory= (boardgamecategory[,1]))

```

Find 5 most popular game categories
```{r}
top_games <- games %>%
  group_by(boardgamecategory) %>%   #group by category
  summarize(n()) %>%                #add count by category
  top_n(5)                        #filter top 5

top_games
```

Filter games from the most popular categories
```{r}
games <- games %>%
  filter(boardgamecategory == top_games$boardgamecategory)

```

## Half-Eye Plot

https://cran.r-project.org/web/packages/ggdist/vignettes/slabinterval.html

Requires ggdist package.
```{r}
library(ggdist)
```

Plot
```{r}
ggplot(games, aes(x = average, y = boardgamecategory)) +        

  stat_halfeye(slab_type="pdf") +  # PDF = probability density function. Can be replaced by histogram or CDF (cumulative density function)

  stat_pointinterval(.width = c(.5, .95)) +     #define the confidence intervals for the bars

  theme_test()  
```

